\documentclass[12pt,letterpaper,oneside,final]{memoir}

% ---------- Options ----------------- %

  % NYU GSAS Requirements
  \usepackage{amsmath}
  \usepackage{amssymb}
  \usepackage{amsthm}
  \usepackage{array}
  \usepackage{bbm}
  \usepackage{booktabs}
  \usepackage{caption}
  \usepackage{color, colortbl}
  \usepackage{dsfont}
  \usepackage{enumitem}
  \usepackage{epstopdf}
  \usepackage[gen]{eurosym}
  \usepackage{float}
  \usepackage[left=1in, right=1in, top=1in, bottom=1in, includefoot, heightrounded,footskip=.75in]{geometry}
  \usepackage[colorlinks=true, citecolor=blue, urlcolor=blue, linkcolor=blue, backref=page]{hyperref}
  \usepackage{kpfonts}
  \usepackage{lipsum}
  \usepackage{longtable}
  \usepackage{makeidx}
  \usepackage{mathpazo}
  \usepackage{mathtools}
  \usepackage{morefloats}
  \usepackage{multirow}
  \usepackage[round]{natbib}
  \usepackage{pifont}
  \usepackage{placeins}
  \usepackage{rotating}
  \usepackage{setspace}
  \usepackage{subcaption}
  \usepackage{tabularx}
  \usepackage{titling}
  \usepackage{tkz-tab}
  \usepackage{units}
  \usepackage{verbatim}
  \usepackage{wasysym}

  % Misc packages
  \usepackage{afterpage}
  \usepackage{amsfonts}
  \usepackage[english]{babel}
  \usepackage{bigdelim}
  \usepackage{epsfig}
  \usepackage{fullpage}
  \usepackage{graphicx}
  \usepackage[utf8]{inputenc}
  \usepackage{latexsym}
  \usepackage{listings}
  \usepackage{pdflscape}
  \usepackage{tgtermes}
  \usepackage{threeparttable}
  \usepackage{tikz}

  % New commands
  \newcommand{\cmark}{\ding{51}} % Nice checkmark for tables
  \newcommand{\xmark}{\ding{55}} % Nice X mark for tables

  % Theorems etc
  \newtheorem{theorem}{Theorem}
  \newtheorem{lemma}[theorem]{Lemma}
  \newtheorem{proposition}[theorem]{Proposition}
  \newtheorem{corollary}[theorem]{Corollary}
  \newtheorem{definition}{Definition}
  \newtheorem{result}{Result}

  % Misc new commands
  \def\citeapos#1{\citeauthor{#1}'s (\citeyear{#1})}
  \newcommand{\var}{\mbox{\textit Var\/}}
  \newcommand{\wh}{\widehat}
  \newcommand{\wt}{\widetilde}
  \newcommand{\ti}{\tilde}
  \newcommand{\TODO}[1]{
    {\color{red} TODO: #1}
  }
  \newcommand{\NOTE}[2]{
    {\color{blue} NOTE (author: #1): #2}
  }

  % Other personalizations
  \setitemize{noitemsep}
  \setenumerate{noitemsep}

  % Footnote without marker
  \newcommand\blfootnote[1]{%
    \begingroup
    \renewcommand\thefootnote{}\footnote{#1}%
    \addtocounter{footnote}{-1}%
    \endgroup
  }

  % Fix position for page numbers
  \pagestyle{plain}

  % Needed to create list of appendices
  % disable everything after the POST hook, but restore after BIB
  \cftinsertcode{POST}{
    \setcounter{tocdepth}{-1}
  }
  \cftinsertcode{BIB}{
    \setcounter{tocdepth}{0}
  }

  \newcommand\tableofcontentsapps{
    \begingroup
      % disable first part
      \cftinsertcode{PRE}{
        \setcounter{tocdepth}{-10}
      }
      % enable down to subsection within appendices
      \cftinsertcode{POST}{
        \setcounter{tocdepth}{1}
      }

      % disable bibliography
      \cftinsertcode{BIB}{
        \setcounter{tocdepth}{-10}
      }

      \renewcommand\contentsname{List of Appendices}
        \tableofcontents
    \endgroup
  }

% OPTIONS END, DOCUMENT STARTS HERE

\begin{document}

\cftinserthook{toc}{PRE}
\fussy
\vbadness=10000 % badness above which bad vboxes are shown. (Default = 10000?)
\frontmatter
\hyphenation{}

\DoubleSpacing

\begin{center}

% ----------
% Title Page
% ----------

\thispagestyle{empty}
\vspace*{15mm} Three Essays in Macroeconomics
\vspace{20mm}
by\\
\vspace{10mm}
Chase G. Coleman \\
\vspace{10mm}
A dissertation submitted in partial fulfillment\\
of the requirements for the degree of\\
Doctor of Philosophy\\
Department of Economics\\
Stern School of Business\\
New York University\\
May, 2019
\end{center}
\vspace{20mm}
\begin{flushright}
{\rule[0pt]{60mm}{0.1mm}}\\ %rule[raise-height]{width}{height} * raise-height specifies how high to raise the rule (optional) * width specifies the length of the rule (mandatory) * height specifies the height of the rule (mandatory)
Laura L. Veldkamp\\
\vspace{7.5mm}
% {\rule[0pt]{60mm}{0.1mm}}\\
\end{flushright}

\newpage

% ---------
% COPYRIGHT
% ---------

%\thispagestyle{empty}

% \begin{center} Copyright © Chase G. Coleman \\
% All Rights Reserved, 2019 \end{center}

%\newpage
%
%\thispagestyle{empty}
%

% ----------
% DEDICATION
% ----------
% \chapter{Dedication}
  %
  % \vspace*{\fill}
  % \begin{center}
  %   \begin{minipage}{.6\textwidth}
  %     I dedicate this dissertation to
  %   \end{minipage}
  % \end{center}
  % \vfill

\newpage

% ---------------
% ACKNOWLEDGMENTS
% ---------------
\chapter{Acknowledgments}

  First and foremost, I am grateful for a truly exceptional committee consisting of Laura Veldkamp,
  Thomas Sargent, and Stan Zin. They have each dedicated a significant portion of their time to
  helping me learn how to study economics and I am a more capable economist because of this
  investment. The innate curiosity shared by the members of my committee inspired me and often
  sparked progress and new insights when I was struggling with how to proceed.

  I am also grateful for the collaborative environment that I was able to work in while here at NYU
  and am lucky to have learned an incredible amount about economics from various students and
  professors, including but not limited to, Felipe Alvarez, Katka Borovickova, Tom Cooley, Richard
  Davies, Fiona Feng, Axelle Ferriere, Sonia Gilbukh, Victoria Gregory, Timothy Hills, Malika
  Krishna, Elliot Lipnowski, Roxana Mihet, Simon Mongey, Jonathan Payne, Alberto Polo, Kim Ruhl,
  Maher Said, Daniel Stackman, Balint Szoke, Venky Venkateswaran, Mike Waugh, Larry White, Peifan
  Wu.

  Dave Backus, though no longer with us, also played a vital role in my development as an economist
  and as a person --- I often miss him and frequently reflect on the many lessons I learned from
  him.

  The Sargent Reading Group served as an effective tool for me to learn about a wide range of
  topics. I am grateful to those who helped organize the reading group while I attended --- Joseba
  Martinez and Balint Szoke --- and am also grateful to each of the particants who took it seriously
  by showing up with interesting papers and by asking insightful, and sometimes difficult,
  questions.

  Outside of NYU, I have also benefitted from a number of wonderful colleagues, including but not
  limited to, Rick Evans, Lilia Maliar, Serguei Maliar, Matt McKay, Kerk Phillips, Jim Savage, Simon
  Scheidegger, John Stachurski, and Natasha Watkins.

  This work is dedicated to my parents, Stephanie and Gregory Coleman, and to my wife, Jessica
  Daniels-Coleman. At different stages of life, they have each supported my desire to learn and
  encouraged me to pursue the paths that this desire has led me to take, often at the price of
  their own convenience.

\newpage

% --------
% ABSTRACT
% --------
\chapter{Abstract}

  \DoubleSpacing

  This dissertation consists of three chapters.

  Chapter 1, `Pareto weights as wedges in two-country models`, was written with Dave Backus, Axelle
  Ferriere, and Spencer Lyon. In this chapter we study how in models with recursive preferences,
  endogenous variation in Pareto weights would be interpreted as wedges from the perspective of a
  frictionless model with additive preferences. We describe the behavior of the relative Pareto
  weight in a two-country world and explore its interaction with consumption and the real exchange
  rate.

  Chapter 2, `Global Solution Methods for Macroeconomic Models`, was written jointly with Spencer
  Lyon, Lilia Maliar, and Serguei Maliar. This chapter first discusses 7 global solution methods in
  the context of a simple neoclassical growth model and then introduces a new novel global solution
  method in the context of a non-trivial New-Keynesian model. We first highlight how algorithm
  choice, even more than programming language, plays an important role in the speed that a model can
  be solved. The new method introduced in this paper is capable of solving a New-Keynesian model
  with an 8-dimensional state-space in mere seconds which makes it possible to perform model
  calibration.

  In chapter 3, `The Cost of Income-Driven Repayment Plans`, I study the implied costs of two
  student loan repayment plans. The first plan, AMR, takes the outstanding loans at the end of
  college and amortizes the payments over 10 years. The second plan, IDR, implements an income
  contingent repayment strategy. I use a detailed model of pre, intra, and post college life to
  compare how these plans differ in enrollment, completion, debt accumulation, and cost to the
  government. I find that while IDR improves overall welfare, that it induces some bad behavior
  during the pre and intra college timeframe which makes it costly for the government to fund such
  a plan.

\newpage

% -----------------
% TABLE OF CONTENTS
% -----------------

  \renewcommand*{\cftappendixname}{Appendix\space}
  \renewcommand*{\contentsname}{Table of Contents}
  \setcounter{tocdepth}{1}% chapters and above
  %this removes TOC listing from TOC
  \begin{KeepFromToc}
  \tableofcontents*
  \end{KeepFromToc}

\clearpage
\newpage
\clearpage

% ---------------
% LIST OF FIGURES
% ---------------

  \listoffigures

\newpage
\clearpage

% --------------
% LIST OF TABLES
% --------------

  \listoftables

\newpage
\clearpage

% ------------------
% LIST OF APPENDICES
% ------------------

  \tableofcontentsapps



\DoubleSpacing
\mainmatter
\clearpage
\cleardoublepage
\phantomsection
%\addcontentsline{toc}{chapter}{Introduction}

% ------------
% Introduction
% ------------
\chapter*[Introduction]{Introduction} \label{ch-intro}

  Introduction

  \setcounter{tocdepth}{1}% chapters and above

% CHAPTER 1
\chapter{Pareto weights as wedges in two-country models}

  \input{sections/BCFL/intro.tex}
  \input{sections/BCFL/two_countries.tex}
  \input{sections/BCFL/calibration.tex}
  \input{sections/BCFL/solution.tex}
  \input{sections/BCFL/equilibrium.tex}
  \input{sections/BCFL/conclusion.tex}

\newpage
\clearpage

% CHAPTER 2
\chapter{Global Solution Methods for Macroeconomic Models}

  \input{sections/CLMM/intro}
  \input{sections/CLMM/languageintro}
  \input{sections/CLMM/growthmodel}
  \input{sections/CLMM/nkmodel}
  \input{sections/CLMM/conclusion}


% CHAPTER 3
\chapter{Student Loans}

Add stuff here

\newpage
\clearpage

%%% APPENDICES
\begin{appendices}

\appendix
\addcontentsline{toc}{chapter}{Appendices}
\cftinserthook{toc}{POST}

% APPENDIX 1
\chapter{Supplementary Material for Chapter 1} \label{sec:Appendix1}

  \input{sections/BCFL/appendix.tex}
  \input{sections/BCFL/tables_and_figs.tex}

% APPENDIX 2
\chapter{Supplementary Material for Chapter 2} \label{sec:Appendix2}

  \input{sections/CLMM/appendix.tex}
  \input{sections/CLMM/tables_and_figs.tex}

\newpage
\clearpage

\end{appendices}

\newpage

\clearpage

\backmatter

%%% BIBLIOGRAPHY

  \cftinserthook{toc}{BIB}
  \begin{OnehalfSpace}
    \bibliographystyle{ecta}
    \bibliography{combined}
  \end{OnehalfSpace}

\end{document}
